---
title: "StatInf_Assignment2_ToothGrowth"
author: "Willianto Asalim"
date: "17/05/2020"
output: html_document
---

The platform specification used:

Spec    | Description
------- | -----------------------
OS      | Windows 10 Pro - 64 bit
CPU     | AMD Ryzen 5 - 3400G
RAM     | 16GB DDR4 3000MHz
Storage | 500GB SSD - M.2 NVMe (PCIe) 
Tool    | RStudio

***



##  Statistical Inference Proj 2: Basic Inferential Data Analysis

<br />


### The Effect Of Vitamin C On Tooth Growth In Guinea Pigs

<br />


### Sypnosis

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

Info on ToothGrowth dataset can be found at the following link: [rdocumentation website](https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/ToothGrowth)


***

```{r setoptions, echo=FALSE}
## Setting Global Option where echo = true so that someone will be able to read the code and results.
knitr::opts_chunk$set(echo = TRUE, results = "asis", tidy = FALSE)
```

#### 1. Load the ToothGrowth data and perform some basic exploratory data analyses

```{r LoadToothGrowth_EDA}
library(datasets) ##Load R Builtin Data sets
data("ToothGrowth") ##Load ToothGrowth Datasets
```

```{r CheckToothGrowthData}
dim(ToothGrowth) ##No. observations and variables
names(ToothGrowth) ##variables names
str(ToothGrowth)

```

<br />

##### The ToothGrowth Dataset contains `r nrow(ToothGrowth)` observations and `r length(ToothGrowth)` variables: `r names(ToothGrowth)`.

<br />

Below is the Exploratory Data Analysis perform for the dataset using Geom Point and Boxplot.
```{r, ToothGrowthScatterPlot}
library(ggplot2) ##Load ggplot2 library

edaPlot <- ggplot(ToothGrowth, aes(factor(dose), len, color = supp))  ##ggplot basic info
edaPlot + geom_point(size = 2) + theme_bw() + labs(x = "Dose", y = "Length", color = "Supplement") +
        scale_color_brewer(palette = "Dark2")


```

***

#### 2. Provide a basic summary of the data.

```{r ToothGrowthSummary}
summary(ToothGrowth)

```

```{r BoxPlot}
edaPlot + geom_boxplot() +
      # facet_grid(.~supp)+
      facet_grid(.~supp, labeller = as_labeller(
            c("OJ" = "Orange juice", 
              "VC" = "Vitamin C"))) +
      labs(title = "Tooth growth of 60 guinea pigs by supplement type based on dosage",
           x = "Dose in milligrams/day", 
           y = "Tooth Lengh") +
      scale_fill_discrete(name = "Dosage of\nvitamin C\nin mg/day")


```


