---
title: "StatInf_Assignment2_ToothGrowth"
author: "Willianto Asalim"
date: "17/05/2020"
output: html_document
---

The platform specification used:

Spec    | Description
------- | -----------------------
OS      | Windows 10 Pro - 64 bit
CPU     | AMD Ryzen 5 - 3400G
RAM     | 16GB DDR4 3000MHz
Storage | 500GB SSD - M.2 NVMe (PCIe) 
Tool    | RStudio

***



##  Statistical Inference Proj 2: Basic Inferential Data Analysis

<br />


### The Effect Of Vitamin C On Tooth Growth In Guinea Pigs

<br />


### Sypnosis

The response is the length of odontoblasts (cells responsible for tooth growth) in 60 guinea pigs. Each animal received one of three dose levels of vitamin C (0.5, 1, and 2 mg/day) by one of two delivery methods, orange juice or ascorbic acid (a form of vitamin C and coded as VC).

Info on ToothGrowth dataset can be found at the following link: [rdocumentation website](https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/ToothGrowth)


***

```{r setoptions, echo=FALSE}
## Setting Global Option where echo = true so that someone will be able to read the code and results.
knitr::opts_chunk$set(echo = TRUE, results = "asis", tidy = FALSE)
```

1. Load the ToothGrowth data and perform some basic exploratory data analyses
```{r LoadToothGrowth_EDA}
library(datasets) 
data("ToothGrowth")
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
